---
title: "Step 01: Downloading RBIEN coordinate data from the server"
author: "Alexa Tyszka"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Initialization
First, we will prepare to download all the data that we might want from a subset of species from the main BIEN database. Open the BIEN package and set up a directory where we will be storing the files.


```{r init, echo=FALSE}
#install.packages("BIEN")
library(BIEN)
setwd("~/atyszka2@uic.edu - Google Drive/Shared drives/walker-lab-shared-drive/preservation_review_ast/04_data_procedure_for_submission/")

```

You may want multiple downloads, each of which should have a separate folder.
```{r dircreate, eval=FALSE}
download.dir.path <- paste(Sys.Date(), "-rbien-bulk-download", sep="")
dir.create(download.dir.path)
setwd(download.dir.path)
getwd()
```

## Get a list of species of interest

You may want to investigate a species or genus of interest. RBIEN provides many tools to investigate such questions. Or you may have a pre-existing vector of species, which you can load here. I'm investigating all species in the database, so here we get a record of all species.
```{r getsp, eval=F}
Blistall<-BIEN_list_all()
```

##Setting up download parameters

```{r dlsetup}
compiled_results <- NULL
n <- 10 #species for each query, and how frequently your .csv file will be saved.
#Lower numbers here will make it easier to troubleshoot but result in many files.
m <- length(Blistall$species) #total number of species
startnum <- 0 #which index we start at, could be useful if server run is interrupted
```


## Running the loop, and final notes
This code was run on a lab computer running Ubuntu 22.04.3 LTS (GNU/Linux 6.2.0-39-generic x86_64), but could also be run for smaller jobs on lighter weight PCs. Adjusting ```m``` above is useful for test-runs, for example only taking 100 species initially. The ```.csv``` files will vary in size and some may be 0 bytes large, but this is just because there were no coordinate records for that species. Records may exist in other forms.

```{r serverrun, eval = FALSE, echo=T}
## try error handling is courtesy of https://stackoverflow.com/questions/31999808/retry-for-loop-r-loop-if-error
for (x in seq(from=startnum, to=m, by=n)) {
  checknum <- 0 # setup for retry error
  message(x, " out of ", m, ", ",x/m ,"%") #print progress
  queries <- na.omit(Blistall$species[x:(x+(n-1))]) #get queries from db
  while(TRUE){ 
    download_try <- try(results <- BIEN_occurrence_species(species = queries,observation.type=T), silent=TRUE)
    if(is(download_try, 'try-error')){
      message("WARNING! timeout on: ", x, "\t retry attempt:", checknum)
      Sys.sleep(20)
      checknum = checknum + 1
    }
    else{
      break
      }
    }
  write.csv(results, file=paste("rbien_coords_dec_14/rbien_dl_from_", x, ".csv", sep=""))
}

```

