---
title: "Downloading RBIEN coordinate data from the server"
author: "Alexa Tyszka"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

## Initialization
First, we will prepare to download all the data that we might want from a subset of species from the main BIEN database. Open the BIEN package and set up a directory where we will be storing the files. For more information on using the BIEN package, please consult [this vignette](http://cran.nexr.com/web/packages/BIEN/vignettes/BIEN_tutorial.html).


```{r init, echo=TRUE}
#install.packages("BIEN")
library(BIEN)
setwd("~/Documents/GitHub/preservation-review-2024/02_download_rbien_sp_list/")

```

You may want multiple downloads, each of which should have a separate folder.
```{r dircreate, eval=FALSE}

download.dir.path <- paste(Sys.Date(), "-rbien-bulk-download", sep="")
dir.create(download.dir.path)
setwd(download.dir.path)
getwd()
```

## Get a list of species of interest

You may want to investigate a species or genus of interest. RBIEN provides many tools to investigate such questions. Or you may have a pre-existing vector of species, which you can load here. I'm investigating all species in the database, so here we get a record of all species.
```{r getsp, eval=T}
Blistall<-BIEN_list_all()
```



##Setting up download parameters

```{r dlsetup}
compiled_results <- NULL
n <- 10 #species for each query, and how frequently your .csv file will be saved.
#Lower numbers here will make it easier to troubleshoot but result in many files.
m <- length(Blistall$species) #total number of species
startnum <- 0 #which index we start at, could be useful if server run is interrupted
```


## Running the loop, and final notes
This code was run on a lab computer running Ubuntu 22.04.3 LTS (GNU/Linux 6.2.0-39-generic x86_64). Adjusting ```m``` is useful for test runs, for example only taking 100 species initially. The ```.csv``` files will vary in size and some may be 0 bytes large, but this is just because there were no coordinate records for that species. Records may exist in other forms.

```{r serverrun, eval = FALSE, echo=T}
## try error handling is courtesy of https://stackoverflow.com/questions/31999808/retry-for-loop-r-loop-if-error
for (x in seq(from=startnum, to=m, by=n)) {
  checknum <- 0 # setup for retry error
  message(x, " out of ", m, ", ",x/m ,"%") #print progress
  queries <- na.omit(Blistall$species[x:(x+(n-1))]) #get queries from db
  while(TRUE){ 
    download_try <- try(results <- BIEN_occurrence_species(species = queries,observation.type=T), silent=TRUE)
    if(is(download_try, 'try-error')){
      message("WARNING! timeout on: ", x, "\t retry attempt:", checknum)
      Sys.sleep(20)
      checknum = checknum + 1
    }
    else{
      break
      }
    }
  write.csv(results, file=paste("rbien_coords_/", x, ".csv", sep=""))
}

```

As an example, we can download the range of all species within the genus *Orzya* present in the BIEN database. In this case, with only a few members of the genus, we can just use the function provided by the BIEN package.

```{r oryza_example, eval=T}
search_genus<-"Oryza"
results <- BIEN_occurrence_genus(genus = search_genus,observation.type=T)
write.csv(results, file=paste(search_genus, ".csv", sep=""))
```

